---
title: "Custom selectors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom selectors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Use a custom selector when none of the bundled rules (`LATEST`, `EARLIEST`, `COUNT`, `COUNT_MORE_THAN_1`, `RANGE_COUNT`, `ALL` -- see `vignette("selector-cookbook")`) reduce a window's concept matches the way you need. This example keeps the highest numeric value recorded in each window.

Unlike a constructor, a selector is not an R function -- it is SQL that `anchor()` runs directly inside its DuckDB connection, where `population_windows` (aliased `w`) and `concepts` (aliased `c`) are already loaded. It must:

- Project exactly these columns: `person_id`, `T0`, `variable_id`, `window_name`, `value`, `date`, `n`.
- Filter on `w.selector = '<NAME>'` so it only claims the metadata rows routed to it.
- Join `concepts AS c` to `population_windows AS w` on `person_id`, `concept_id`, and `c.date BETWEEN w.window_start AND w.window_end`, the same join every bundled template uses.

Selector metadata value `MAX_VALUE` resolves to an object named `max_value_selector`, built with `make_selector()`.

```{r selector}
max_value_selector <- make_selector("
  SELECT
      w.person_id,
      w.T0,
      w.variable_id,
      w.window_name,
      CAST(MAX(TRY_CAST(c.value AS DOUBLE)) AS VARCHAR) AS value,
      MAX(c.date) AS date,
      COUNT(*) AS n
  FROM population_windows AS w
  INNER JOIN concepts AS c
      ON c.person_id = w.person_id
     AND c.concept_id = w.concept_id
     AND c.date BETWEEN w.window_start AND w.window_end
  WHERE w.selector = 'MAX_VALUE'
  GROUP BY w.person_id, w.T0, w.variable_id, w.window_name
")
```

`make_selector()` only checks that `selector_query` is a non-empty string and that it selects every required column by name -- it cannot validate the SQL itself, since the tables it references only exist once `anchor()` opens its DuckDB connection. A typo surfaces as a database error the first time the selector actually runs.

```{r inputs}
population <- data.table(
  person_id = c("1", "2", "3"),
  T0 = as.Date(c("2024-01-01", "2024-01-15", "2024-02-01"))
)
metadata <- data.table(
  variable_id = "highest_lab_value",
  concept_id = "LAB_X",
  constructor = "GENERIC",
  selector = "MAX_VALUE",
  start_offset = -90L,
  end_offset = 0L
)
concepts <- data.table(
  person_id = c("1", "1", "2"),
  concept_id = "LAB_X",
  date = as.Date(c("2023-11-01", "2023-12-15", "2023-12-20")),
  value = c("4.2", "7.8", "3.1")
)
```

```{r anchor}
hive <- tempfile("selector-hive-")

# anchor() resolves custom selectors from the global environment, the same
# way it resolves custom constructors.
had_existing_selector <- exists(
  "max_value_selector", envir = globalenv(), inherits = FALSE
)
if (had_existing_selector) {
  existing_selector <- get(
    "max_value_selector", envir = globalenv(), inherits = FALSE
  )
}
assign("max_value_selector", max_value_selector, envir = globalenv())

anchor(population, metadata, concepts, anchor_hive_path = hive)
get_anchor_result(metadata, hive, result_shape = "long")[
  , .(person_id, value, date)
]
```

Person 1 gets `7.8`, the higher of two matches; person 2 gets `3.1`, its only match; person 3 has no `LAB_X` records at all, so it produces no row -- the same "silent gap, not an error" behavior every selector has when nothing matches.

Built-in and custom selector rows may coexist in one metadata table, and a metadata file can reference a custom selector without erroring during validation: `metadata_supported_selectors()` accepts any selector name that either ships with anchoR or resolves via `selector_env` (default `globalenv()`), the same two-tier lookup `resolve_window_constructor()` uses for constructors.

```{r cleanup, include=FALSE}
unlink(hive, recursive = TRUE)
if (had_existing_selector) {
  assign("max_value_selector", existing_selector, envir = globalenv())
} else {
  rm("max_value_selector", envir = globalenv())
}
```
