---
title: "Custom window constructors"
vignette: >
  %\VignetteIndexEntry{Custom window constructors}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Use a custom constructor when window boundaries cannot be expressed by the built-in generic or episode rules. This example starts at an index date and ends at the earlier of recorded follow-up or 180 days after index.

Constructor metadata value `CAPPED_FOLLOWUP` resolves to a function named `capped_followup_window`. The function must return its input rows with `window_start` and `window_end` columns.

```{r constructor}
capped_followup_window <- make_constructor(
  transform_fn = function(window_dt) {
    window_dt[, `:=`(
      window_start = as.Date(NA),
      window_end = as.Date(NA),
      .follow_up_cap = as.Date(NA)
    )]

    for (col in unique(window_dt$anchor_start_col)) {
      window_dt[
        anchor_start_col == col,
        `:=`(
          window_start = as.Date(get(col) + start_offset),
          .follow_up_cap = as.Date(get(col) + 180L)
        )
      ]
    }
    for (col in unique(window_dt$anchor_end_col)) {
      window_dt[
        anchor_end_col == col,
        window_end := as.Date(pmin(get(col) + end_offset, .follow_up_cap))
      ]
    }

    window_dt[, .follow_up_cap := NULL]
    window_dt[]
  },
  required_cols = c(
    "anchor_start_col", "anchor_end_col", "start_offset", "end_offset"
  )
)
```

The constructor uses canonical metadata fields so validation and the anchoring pipeline retain everything it needs.

```{r inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", "2024-01-01")),
  follow_up_end = as.Date(c("2024-03-01", "2025-01-01"))
)
metadata <- data.table(
  variable_id = "first_follow_up_event",
  concept_id = "EVENT",
  constructor = "CAPPED_FOLLOWUP",
  selector = "EARLIEST",
  start_offset = 0L,
  end_offset = 0L,
  anchor_start_col = "T0",
  anchor_end_col = "follow_up_end"
)

define_window(
  population, metadata,
  constructor_env = environment()
)[
  , .(person_id, window_start, window_end, window_valid)
]
```

Person 1 is capped by recorded follow-up; person 2 is capped at `T0 + 180`. Test those boundaries with `define_window()` before applying concepts.

```{r anchor}
concepts <- data.table(
  person_id = c("1", "2", "2"),
  concept_id = "EVENT",
  date = as.Date(c("2024-02-01", "2024-05-01", "2024-08-01")),
  value = "TRUE"
)
hive <- tempfile("custom-hive-")

# anchor() resolves custom constructors from the global environment.
had_existing_constructor <- exists(
  "capped_followup_window", envir = globalenv(), inherits = FALSE
)
if (had_existing_constructor) {
  existing_constructor <- get(
    "capped_followup_window", envir = globalenv(), inherits = FALSE
  )
}
assign(
  "capped_followup_window", capped_followup_window,
  envir = globalenv()
)
anchor(population, metadata, concepts, anchor_hive_path = hive)
get_anchor_result(metadata, hive, result_shape = "long")
```

Custom constructors are discovered in the global environment by the standard `anchor()` workflow. `define_window()` additionally accepts `constructor_env` for explicit lookup during isolated testing. Built-in and custom constructor rows may coexist in one metadata table.

Current normalization retains only anchoR's canonical metadata fields. Encode custom boundaries using canonical anchor/offset columns or implement preprocessing upstream; arbitrary extra metadata columns are not passed through validation.

```{r cleanup, include=FALSE}
unlink(hive, recursive = TRUE)
if (had_existing_constructor) {
  assign(
    "capped_followup_window", existing_constructor,
    envir = globalenv()
  )
} else {
  rm("capped_followup_window", envir = globalenv())
}
```
