---
title: "Episode-based windows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Episode-based windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Episode-based constructors build windows from repeatable start/end periods. The built-in names use pregnancy terminology, but the underlying representation is generic.

Episodes live in their own long table, one row per episode, with `person_id`, `start_episode`, and `end_episode` columns. It's passed as a separate `episodes` argument, alongside `population` and `metadata`; anchoR nests it onto `population` internally.

```{r population}
episodes <- data.table(
    person_id = "1",
    start_episode = as.Date(c("2023-01-01", "2024-03-01", "2025-11-01")),
    end_episode = as.Date(c("2023-09-01", "2024-11-15", "2026-08-01"))
)

population <- data.table(
    person_id = "1",
    T0 = as.Date("2026-02-15")
)
```


`T0` falls inside the third episode, so that one is "current" and the other two are "prior". The four constructors select the current episode, every prior episode, both combined, or the gaps outside every episode.

```{r metadata}
metadata <- data.table(
    variable_id = c("current", "prior", "current_and_prior", "outside"),
    concept_id = "DEMO",
    constructor = c(
        "in_current_pregnancy", "in_prior_pregnancy",
        "in_current_and_prior", "outside_all_pregnancy"
    ),
    selector = "ALL",
    start_offset = 0L, # unused by episode-based constructors, still required
    end_offset = 0L,
    anchor_start_offset = c(NA_real_, NA_real_, NA_real_, -1172L),
    anchor_end_offset = c(NA_real_, NA_real_, NA_real_, 0L),
    before_start_episode_offset = c(0, 0, 0, NA_real_),
    after_start_episode_offset = NA_real_,
    before_end_episode_offset = NA_real_,
    after_end_episode_offset = c(14, 0, 30, NA_real_)
)

define_window(population, metadata, episodes = episodes)[
    window_valid == TRUE,
    .(variable_id, window_start, window_end)
]
```


Every episode-based window is built from two independent border pairs: `before_start_episode_offset`/`after_start_episode_offset` (relative to the episode's own `start_episode`) and `before_end_episode_offset`/`after_end_episode_offset` (relative to `end_episode`). Each pair is read as `edge + offset`, `NA` meaning that side isn't set. `before_start_episode_offset` and `after_end_episode_offset` are the two "outer" sides (they point away from the episode); `after_start_episode_offset` and `before_end_episode_offset` are the "inner" sides (they point into it):

- **Either outer side set alone** (as above: `current`'s `before_start_episode_offset = 0` and `after_end_episode_offset = 14`): the whole window becomes one shared window, that side's offset defining its edge directly, the other pair contributing whichever single value it has (or the unshifted edge, if it has nothing) for the opposite edge. `current`'s start is pinned to the episode's own start (`before_start_episode_offset = 0`); its end is 14 days past the episode's own end (`after_end_episode_offset = 14`).
- **Only an inner side set, with no outer side set anywhere**: each such pair becomes its own self-contained region instead, `[edge + before_offset, edge + after_offset]` (missing side defaulting to `0`), rather than joining a shared window.
- **Both sides of a pair set**: that pair is always its own self-contained region, regardless of the other pair.

See `documentation/definitions/Episode-Based Window Engine.md` for the complete rule, including what happens when the two pairs disagree (one outer, one inner).

`outside_all_pregnancy` works differently: it doesn't use the before/after episode offsets at all, only `anchor_start_offset`/`anchor_end_offset`, which define the overall search range (`[T0 - 1172, T0]` above) that the gaps between episodes are computed within.

The border-offset formula has no idea how long a selected episode actually lasted, though — `after_start_episode_offset = 400` always adds 400 days to `start_episode`, whether or not that episode's own `end_episode` came sooner. Two optional logical columns, `cap_start_to_episode`/`cap_end_to_episode` (both `NA`/unset by default, not read by `outside_all_pregnancy`), pull a window back inside its own selected episode's real bounds when that happens:

```{r episode-cap}
capped_metadata <- data.table(
    variable_id = "current_capped",
    concept_id = "DEMO",
    constructor = "in_current_pregnancy",
    selector = "ALL",
    start_offset = 0L,
    end_offset = 0L,
    after_start_episode_offset = 400L, # 400 days after episode C's own start
    cap_end_to_episode = TRUE # ... but never past episode C's own end
)

define_window(population, capped_metadata, episodes = episodes)[
    window_valid == TRUE,
    .(variable_id, window_start, window_end)
]
```


Without the cap, `after_start_episode_offset = 400` alone would run to `2026-12-06`, four months past episode C's own end (`2026-08-01`); with `cap_end_to_episode = TRUE`, the window stops at the episode's real end instead.

`anchor_start_offset`/`anchor_end_offset` aren't only `outside_all_pregnancy`'s search range, though — they're also a hard boundary applied to *every* constructor's window: `[T0 + anchor_start_offset, T0 + anchor_end_offset]`, clipped in independently of the border-offset formula and the episode-bounds cap above. Left `NA` (as for `current`/`prior`/`current_and_prior` in the metadata above), that boundary is simply not enforced. Setting it clips whatever window the border-offset formula produced, even down to nothing:

```{r anchor-clip}
clipped_metadata <- data.table(
    variable_id = "current_clipped",
    concept_id = "DEMO",
    constructor = "in_current_pregnancy",
    selector = "ALL",
    start_offset = 0L,
    end_offset = 0L,
    anchor_start_offset = -30L, # boundary: 30 days before T0 ...
    anchor_end_offset = 0L, # ... through T0 itself
    before_start_episode_offset = NA_real_,
    after_start_episode_offset = NA_real_,
    before_end_episode_offset = NA_real_,
    after_end_episode_offset = NA_real_
)

define_window(population, clipped_metadata, episodes = episodes)[
    window_valid == TRUE,
    .(variable_id, window_start, window_end)
]
```


The current episode's own unshifted span (`[2025-11-01, 2026-08-01]`) gets clipped down to `[2026-01-16, 2026-02-15]`, the last 30 days before `T0`.

The internal engine can produce several candidate windows for one metadata row (`prior` and `current_and_prior` both do, one per selected episode). Selectors aggregate matches from those windows into the same output key. Candidate windows are not deduplicated: if they overlap, one concept event can join more than once and affect `COUNT` or `ALL`. Keep episode windows non-overlapping when distinct-event counts are required.

For applications unrelated to pregnancy, either reuse these constructors (the underlying episodes can be any repeatable period), or create project-specific names with `make_constructor()` when the pregnancy-oriented vocabulary would be confusing.
