---
title: "Episode-based windows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Episode-based windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Episode-based constructors build windows from repeatable start/end periods. The built-in names use pregnancy terminology, but the underlying representation is generic.

Each population row holds a list-column whose element is a table with `event_start` and `event_end`.

```{r population}
episodes <- data.table(
  event_start = as.Date(c("2023-01-01", "2024-03-01", "2025-11-01")),
  event_end = as.Date(c("2023-09-01", "2024-11-15", "2026-08-01"))
)

population <- data.table(
  person_id = "1",
  T0 = as.Date("2026-02-15"),
  pregnancy_episodes = list(episodes)
)
```

The four constructors select prior episodes, the current episode through the anchor, the full current episode, or gaps outside every episode.

```{r metadata}
metadata <- data.table(
  variable_id = c("prior", "since_start", "anytime", "outside"),
  concept_id = "DEMO",
  constructor = c(
    "IN_PRIOR_PREG", "SINCE_START_CURRENT_PREG",
    "ANYTIME_CURRENT_PREG", "OUTSIDE_ALL_PREG"
  ),
  selector = "ALL",
  start_offset = c(0L, 0L, 0L, -1172L),
  end_offset = c(30L, 0L, 14L, 0L),
  event_col = "pregnancy_episodes"
)

define_window(population, metadata)[
  window_valid == TRUE,
  .(variable_id, window_start, window_end)
]
```

`IN_PRIOR_PREG` can optionally cap each window at `event_start + end_cap_offset`. Its separate `start_look_back` and `end_look_back` fields filter which prior episodes are eligible; they do not clip a surviving episode's window. `OUTSIDE_ALL_PREG` instead uses `start_offset` and `end_offset` as the overall anchor-relative search range.

The internal engine can produce several candidate windows for one metadata row. Selectors aggregate matches from those windows into the same output key. Candidate windows are not deduplicated: if they overlap, one concept event can join more than once and affect `COUNT` or `ALL`. Keep episode windows non-overlapping when distinct-event counts are required.

For applications unrelated to pregnancy, either reuse these constructors with a clearly documented `event_col`, or create project-specific names with `make_constructor()` when the pregnancy-oriented vocabulary would be confusing.
