---
title: "Episode-based windows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Episode-based windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Episode-based constructors build windows from repeatable start/end periods. The built-in names use pregnancy terminology, but the underlying representation is generic.

Episodes live in their own long table, one row per episode, with `person_id`, `start_episode`, and `end_episode` columns. It's passed as a separate `episodes` argument, alongside `population` and `metadata`; anchoR nests it onto `population` internally.

```{r population}
episodes <- data.table(
    person_id = "1",
    start_episode = as.Date(c("2023-01-01", "2024-03-01", "2025-11-01")),
    end_episode = as.Date(c("2023-09-01", "2024-11-15", "2026-08-01"))
)

population <- data.table(
    person_id = "1",
    T0 = as.Date("2026-02-15")
)
```

`T0` falls inside the third episode, so that one is "current" and the other two are "prior". The four constructors select the current episode, every prior episode, both combined, or the gaps outside every episode.

```{r metadata}
metadata <- data.table(
    variable_id = c("current", "prior", "current_and_prior", "outside"),
    concept_id = "DEMO",
    constructor = c(
        "in_current_pregnancy", "in_prior_pregnancy",
        "in_current_and_prior", "outside_all_pregnancy"
    ),
    selector = "ALL",
    start_offset = 0L, # unused by episode-based constructors, still required
    end_offset = 0L,
    anchor_start_offset = c(NA_real_, NA_real_, NA_real_, -1172L),
    anchor_end_offset = c(NA_real_, NA_real_, NA_real_, 0L),
    before_start_episode_offset = c(0, 0, 0, NA_real_),
    after_start_episode_offset = NA_real_,
    before_end_episode_offset = NA_real_,
    after_end_episode_offset = c(14, 0, 30, NA_real_)
)

define_window(population, metadata, episodes = episodes)[
    window_valid == TRUE,
    .(variable_id, window_start, window_end)
]
```

Every episode-based window is built from two independent border pairs: `before_start_episode_offset`/`after_start_episode_offset` (relative to the episode's own `start_episode`) and `before_end_episode_offset`/`after_end_episode_offset` (relative to `end_episode`). Each pair is read as `edge + offset`, `NA` meaning that side isn't set:

- **At most one side of a pair set** (as above, every pair here has only one non-`NA` side): that edge contributes `edge + offset` to a single shared window. `current`'s `before_start_episode_offset = 0` pins the window's start to the episode's own start; `after_end_episode_offset = 14` extends the end 14 days past the episode's own end.
- **Both sides of a pair set**: that pair stops contributing to the shared window and instead becomes its own self-contained region near that edge, `[edge + before_offset, edge + after_offset]` see `documentation/definitions/Episode-Based Window Engine.md` for the full rule, including what happens when only one pair is fully specified.

`outside_all_pregnancy` works differently: it doesn't use the before/after episode offsets at all, only `anchor_start_offset`/`anchor_end_offset`, which define the overall search range (`[T0 - 1172, T0]` above) that the gaps between episodes are computed within.

The internal engine can produce several candidate windows for one metadata row (`prior` and `current_and_prior` both do, one per selected episode). Selectors aggregate matches from those windows into the same output key. Candidate windows are not deduplicated: if they overlap, one concept event can join more than once and affect `COUNT` or `ALL`. Keep episode windows non-overlapping when distinct-event counts are required.

For applications unrelated to pregnancy, either reuse these constructors (the underlying episodes can be any repeatable period), or create project-specific names with `make_constructor()` when the pregnancy-oriented vocabulary would be confusing.
