---
title: "Completing and imputing wide results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Completing and imputing wide results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Imputation occurs during wide retrieval, after population keys and requested result columns have been completed. It does not modify persisted parquet or long output.

```{r inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", "2024-01-01")),
  group = c("EXPOSED", "CONTROL")
)
metadata <- data.table(
  variable_id = c("diagnosis", "category", "expected_unknown"),
  concept_id = c("DIAGNOSIS", "CATEGORY", "UNKNOWN"),
  constructor = "GENERIC",
  selector = "LATEST",
  window_name = "lookback",
  start_offset = -365L,
  end_offset = 0L,
  is_expected_missing = c(FALSE, FALSE, TRUE),
  variable_type = c("BOOL", "CAT", "BOOL")
)
concepts <- data.table(
  person_id = "1",
  concept_id = "DIAGNOSIS",
  date = as.Date("2023-10-01"),
  value = "TRUE"
)
hive <- tempfile("imputation-hive-")
anchor(population, metadata, concepts, anchor_hive_path = hive)
```

Without imputation, completed cells are `NA`.

```{r without-imputation}
get_anchor_result(
  metadata, hive, population = population,
  result_shape = "wide"
)
```

```{r with-imputation}
imputed <- get_anchor_result(
  metadata, hive, population = population,
  result_shape = "wide",
  impute_missing = TRUE
)
imputed
```

Rules are intentionally narrow:

- `BOOL`, `BOOLEAN`, `LOGICAL`, and `TF` become logical and missing cells become `FALSE`.
- `CAT` and `FACTOR` missing cells become `0`.
- `is_expected_missing = TRUE` prevents imputation for that variable.
- Date columns and unsupported variable types remain missing.
- If required imputation metadata is partly supplied, anchoR warns and skips imputation.

These defaults encode domain assumptions, not generic missing-data methodology. Use them only when a missing concept record means false or the categorical zero level, not just "unknown." Otherwise keep `NA` and handle missingness downstream.

Imputation currently targets columns named `value_<variable_id>`. When `cast_window = TRUE` produces `value_<window_name>_<variable_id>`, those columns are not matched by the current imputation helper; impute them downstream or use the default non-window-cast shape.

```{r cleanup, include=FALSE}
unlink(hive, recursive = TRUE)
```
