---
title: "Migrating legacy metadata"
vignette: >
  %\VignetteIndexEntry{Migrating legacy metadata}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Older study-variable files often use BRIDGE-oriented names. anchoR automatically recognizes only three aliases: `date_extraction_func`, `anchor_date_start`, and `anchor_date_end`. Window names, offsets, and constructors must be renamed explicitly.

```{r legacy}
legacy <- data.table(
  variable_id = c("recent_vaccine", "unsupported_variable"),
  concept_id = c("VACCINE", "OTHER"),
  window = c("lookback", "lookback"),
  start = c(-365, -30),
  end = c(0, 0),
  window_definition = c("GENERIC", "GENERIC"),
  date_extraction_func = c("latest", "not implemented"),
  anchor_date_start = c("T0", "T0"),
  anchor_date_end = c("T0", "T0")
)

setnames(
  legacy,
  old = c("window", "start", "end", "window_definition"),
  new = c("window_name", "start_offset", "end_offset", "constructor"),
  skip_absent = TRUE
)
```

Filter unsupported rows only when dropping them is an explicit pipeline decision.

```{r filter}
migrated <- suppressWarnings(filter_supported_metadata(legacy))
migrated[, .(
  variable_id, window_name, start_offset, end_offset,
  date_extraction_func
)]
```

Validation applies the automatic aliases and canonical types.

```{r validate}
population <- data.table(person_id = "1", T0 = as.Date("2024-01-01"))
normalized <- validate_anchor_inputs(population, migrated)$metadata
normalized[, .(
  variable_id, selector, constructor, window_name,
  anchor_start_col, anchor_end_col, start_offset, end_offset
)]
```

Before migration, retain an audit table containing every dropped row and its original selector. Do not silently reinterpret old `start_look_back`/`end_look_back` fields as generic offsets: in current anchoR those names are reserved for the `IN_PRIOR_PREG` episode-eligibility filter.
