---
title: "Migrating legacy metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Migrating legacy metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Older study-variable files often use BRIDGE-oriented names. anchoR automatically recognizes only three aliases: `date_extraction_func`, `anchor_date_start`, and `anchor_date_end`. Window names, offsets, and constructors must be renamed explicitly.

```{r legacy}
legacy <- data.table(
    variable_id = c("recent_vaccine", "unsupported_variable"),
    concept_id = c("VACCINE", "OTHER"),
    window = c("lookback", "lookback"),
    start = c(-365, -30),
    end = c(0, 0),
    window_definition = c("GENERIC", "GENERIC"),
    date_extraction_func = c("latest", "not implemented"),
    anchor_date_start = c("T0", "T0"),
    anchor_date_end = c("T0", "T0")
)

setnames(
    legacy,
    old = c("window", "start", "end", "window_definition"),
    new = c("window_name", "start_offset", "end_offset", "constructor"),
    skip_absent = TRUE
)
```

Filter unsupported rows only when dropping them is an explicit pipeline decision.

```{r filter}
migrated <- suppressWarnings(filter_supported_metadata(legacy))
migrated[, .(
    variable_id, window_name, start_offset, end_offset,
    date_extraction_func
)]
```

Validation applies the automatic aliases and canonical types.

```{r validate}
population <- data.table(person_id = "1", T0 = as.Date("2024-01-01"))
normalized <- validate_anchor_inputs(population, migrated)$metadata
normalized[, .(
    variable_id, selector, constructor, window_name,
    anchor_start_col, anchor_end_col, start_offset, end_offset
)]
```

Before migration, retain an audit table containing every dropped row and its original selector. `start_look_back`/`end_look_back` no longer exist as metadata columns in current anchoR; episode-based metadata now uses `anchor_start_offset`/`anchor_end_offset` and the four `before_*_episode_offset`/`after_*_episode_offset` columns instead (see `vignette("episode-windows", package = "anchoR")`) do not silently reinterpret old `start_look_back`/`end_look_back` values as any of those, since the semantics changed.
