---
title: "Designing multiple windows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing multiple windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

A variable can have several clinically meaningful windows. Repeat `variable_id`, assign a unique `window_name`, and choose offsets and selectors independently.

```{r inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", "2024-01-01"))
)
metadata <- data.table(
  variable_id = "outcome",
  concept_id = "OUTCOME",
  constructor = "GENERIC",
  selector = "EARLIEST",
  window_name = c("lookback", "index", "risk", "control"),
  start_offset = c(-365L, 0L, 1L, 43L),
  end_offset = c(-1L, 0L, 42L, 84L)
)
concepts <- data.table(
  person_id = c("1", "1", "1", "1"),
  concept_id = "OUTCOME",
  date = as.Date(c("2023-12-20", "2024-01-01", "2024-01-20", "2024-03-01")),
  value = "TRUE"
)
```

Inspecting windows makes gaps and boundary choices explicit.

```{r windows}
define_window(population, metadata)[
  person_id == "1",
  .(window_name, window_start, window_end)
]
```

```{r outputs}
hive <- tempfile("window-hive-")
anchor(population, metadata, concepts, anchor_hive_path = hive)

get_anchor_result(metadata, hive, result_shape = "long")
get_anchor_result(
  metadata, hive, population = population,
  result_shape = "wide"
)
get_anchor_result(
  metadata, hive, population = population,
  result_shape = "wide", cast_window = TRUE
)
```

Metadata can request a subset after anchoring. Supplying one explicit `window_name` restricts retrieval to that variable/window pair; an `NA` window name acts as a wildcard for that variable.

```{r subset}
get_anchor_result(
  metadata[window_name == "risk"],
  hive,
  result_shape = "long"
)
```

Use the default wide form when windows are naturally rows and `cast_window = TRUE` when downstream modeling requires one row per population key.

```{r cleanup, include=FALSE}
unlink(hive, recursive = TRUE)
```
