---
title: "Parquet and DuckDB concept sources"
vignette: >
  %\VignetteIndexEntry{Parquet and DuckDB concept sources}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Production concept stores can remain on disk. anchoR queries parquet directly and attaches DuckDB databases read-only, filtering to metadata concept IDs before selector execution.

```{r inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", "2024-01-01")),
  group = c("EXPOSED", "CONTROL")
)
metadata <- data.table(
  variable_id = c("recent_vaccine", "hospital_count"),
  concept_id = c("VACCINE", "HOSP"),
  constructor = "GENERIC",
  selector = c("LATEST", "COUNT"),
  start_offset = -365L,
  end_offset = 0L
)
concepts <- data.table(
  person_id = c("1", "2", "2"),
  concept_id = c("VACCINE", "HOSP", "HOSP"),
  date = as.Date(c("2023-10-01", "2023-08-01", "2023-11-01")),
  value = "TRUE"
)
```

## Hive-partitioned parquet

```{r parquet}
parquet_dir <- tempfile("concept-parquet-")
dir.create(parquet_dir)
con <- DBI::dbConnect(duckdb::duckdb(), dbdir = ":memory:")
DBI::dbWriteTable(con, "concept_table", concepts)
quoted_parquet <- as.character(DBI::dbQuoteString(con, parquet_dir))
DBI::dbExecute(
  con,
  sprintf(
    "COPY concept_table TO %s (FORMAT PARQUET, PARTITION_BY (concept_id))",
    quoted_parquet
  )
)
DBI::dbDisconnect(con, shutdown = TRUE)

hive <- tempfile("production-hive-")
anchor_by_variable(
  population, metadata, parquet_dir,
  anchor_hive_path = hive,
  chunk_size = 1L
)
get_anchor_result(
  metadata, hive, population = population,
  result_shape = "wide", cast_window = TRUE
)
```

Directories, individual parquet files, vectors of files, and glob patterns are accepted. Hive partitioning by `concept_id` lets DuckDB prune irrelevant partitions.

## DuckDB database

The database must contain a table named `concept_table` with `person_id`, `concept_id`, `date`, and `value`.

```{r duckdb}
duckdb_path <- tempfile(fileext = ".duckdb")
updated_concepts <- rbind(
  concepts,
  data.table(
    person_id = "1", concept_id = "VACCINE",
    date = as.Date("2023-12-15"), value = "TRUE"
  )
)
con <- DBI::dbConnect(duckdb::duckdb(), dbdir = duckdb_path)
DBI::dbWriteTable(con, "concept_table", updated_concepts)
DBI::dbDisconnect(con, shutdown = TRUE)

anchor_by_variable(
  population,
  metadata[variable_id == "recent_vaccine"],
  duckdb_path,
  anchor_hive_path = hive,
  chunk_size = 1L
)
get_anchor_result(metadata, hive, result_shape = "long")
```

The second run replaces only `recent_vaccine`; the existing `hospital_count` partition remains. A recomputed variable with no matches has its previous partition removed. If that leaves the entire hive without parquet files, check for the empty directory before retrieval.

Choose batching based on scale:

- `anchor()` minimizes orchestration and handles all requested variables together.
- `anchor_by_variable()` bounds working-set and replacement scope with `chunk_size`.
- `anchor_by_selector()` minimizes selector scans by grouping all variables using each selector.

```{r cleanup, include=FALSE}
unlink(c(parquet_dir, duckdb_path, hive), recursive = TRUE)
```
