---
title: "Selector cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Selector cookbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

Selectors reduce concept matches after inclusive window filtering. This example applies every bundled selector to the same person.

```{r inputs}
population <- data.table(person_id = "1", T0 = as.Date("2024-01-10"))
selectors <- available_selectors()
metadata <- data.table(
  variable_id = paste0("example_", tolower(selectors)),
  concept_id = "MEASURE",
  constructor = "GENERIC",
  selector = selectors,
  start_offset = -30L,
  end_offset = 0L,
  range_min = 2,
  range_max = 10
)
concepts <- data.table(
  person_id = "1",
  concept_id = "MEASURE",
  date = as.Date(c("2024-01-01", "2024-01-05", "2024-01-05")),
  value = c("1", "9", "10")
)

hive <- tempfile("selector-hive-")
anchor(population, metadata, concepts, anchor_hive_path = hive)
result <- get_anchor_result(metadata, hive, result_shape = "long")
result[, .(variable_id, date, value)]
```

Important details:

- `LATEST` and `EARLIEST` return one row. Date ties use the lexicographically largest normalized character value, so `"9"` sorts after `"10"`.
- `COUNT` persists only when at least one record matches; it does not write zero rows.
- `COUNT_MORE_THAN_1` writes `TRUE` only for two or more joined matches.
- `RANGE_COUNT` uses inclusive numeric bounds and ignores values that cannot be converted to numbers.
- `ALL` returns every joined match and normally requires long output.
- Overlapping candidate windows can join the same concept record multiple times.

Selector output `value` is character in persisted/public long results. Convert it to the type you need after retrieval rather than assuming the source type survives every SQL selector.

```{r cleanup, include=FALSE}
unlink(hive, recursive = TRUE)
```
