---
title: "Standard anchoring windows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standard anchoring windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

anchoR combines a population, variable metadata, and concept events. A generic window is an inclusive date range obtained by adding day offsets to an anchor date.

```{r inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", "2024-01-01")),
  group = c("EXPOSED", "CONTROL")
)

metadata <- data.table(
  variable_id = "recent_flu_vaccine",
  concept_id = "FLU_VAX",
  constructor = "GENERIC",
  selector = "LATEST",
  window_name = "one_year",
  start_offset = -365L,
  end_offset = 0L
)

concepts <- data.table(
  person_id = "1",
  concept_id = "FLU_VAX",
  date = as.Date("2023-10-01"),
  value = "TRUE"
)
```

`define_window()` is useful for inspecting the design before reading any concepts.

```{r windows}
define_window(population, metadata)[
  , .(person_id, T0, variable_id, window_name, window_start, window_end)
]
```

`anchor()` persists sparse matches in a parquet Hive. Long output contains only matches.

```{r anchor}
hive_path <- tempfile("anchor-hive-")
anchor(population, metadata, concepts, anchor_hive_path = hive_path)

get_anchor_result(metadata, hive_path, result_shape = "long")
```

Supplying the population to wide output completes missing population keys and reattaches its additional columns.

```{r wide}
get_anchor_result(
  metadata,
  hive_path,
  population = population,
  result_shape = "wide"
)
```

For multiple windows, repeat `variable_id` with distinct `window_name` values. The default wide form retains one row per `person_id`, `T0`, and `window_name`; `cast_window = TRUE` folds the window into each result column name. Use long output for `ALL`, which can put several records into one wide cell. `LATEST` and `EARLIEST` break date ties with the lexicographically largest normalized character value.

Available selectors are returned by `available_selectors()`. Their complete semantics, including inclusive boundaries and tie behavior, are documented in the selector reference on the package website.

```{r cleanup, include=FALSE}
unlink(hive_path, recursive = TRUE)
```
