---
title: "Troubleshooting and validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Troubleshooting and validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(anchoR)
library(data.table)
```

## Inspect windows before anchoring

Most problems are easier to diagnose before concepts and DuckDB are involved. Start with `validate_anchor_inputs()` and `define_window()`.

```{r valid-inputs}
population <- data.table(
  person_id = c("1", "2"),
  T0 = as.Date(c("2024-01-01", NA))
)
metadata <- data.table(
  variable_id = "recent_event",
  concept_id = "EVENT",
  constructor = "GENERIC",
  selector = "LATEST",
  start_offset = -30L,
  end_offset = 0L
)

validated <- validate_anchor_inputs(population, metadata)
windows <- define_window(population, metadata)
windows[, .(person_id, window_start, window_end, window_valid)]
```

A missing anchor creates an invalid window. `anchor()` sends only valid windows to selector SQL.

## Frequent failures

| symptom | likely cause | resolution |
| --- | --- | --- |
| Missing anchor-column error | Metadata references a population field that is absent | Correct `anchor_start_col`/`anchor_end_col` or add the field |
| Date-format error | Primary anchor is not `Date` or strict `YYYY-mm-dd` text | Parse with `as.Date()` before anchoring |
| Unsupported selector | Typo, missing selector, or no bundled SQL template | Check `available_selectors()`; optionally use `filter_supported_metadata()` |
| Window function does not exist | Custom constructor name does not resolve to `<name>_window` | Define it in `constructor_env` and use the required name |
| No parquet files found | Empty/wrong concepts directory | Check the path and produce at least one parquet file |
| Wide output is ambiguous | More than one result occupies a wide cell, commonly from `ALL` | Request long output or reduce events upstream |
| Duplicate population warning | Rows share `person_id/T0` but disagree on extra fields | Choose a stable unique population key upstream; anchoR keeps the first row |
| Empty-hive DuckDB error | No variable produced a parquet partition | Detect an empty hive before calling `get_anchor_result()` |

## Unsupported metadata

`anchor()` stops rather than silently dropping study variables. Inspect deliberate filtering separately:

```{r selector-filter}
candidate_metadata <- rbind(
  metadata,
  copy(metadata)[, `:=`(variable_id = "bad", selector = "NOT_IMPLEMENTED")]
)

available_selectors()
supported_metadata <- suppressWarnings(
  filter_supported_metadata(candidate_metadata)
)
supported_metadata[, .(variable_id, selector)]
```

## Episode checks

Episode list-columns must contain `event_start` and `event_end`. Invalid or overlapping periods should be corrected upstream. Multiple candidate windows are joined independently; overlapping windows can cause one concept event to contribute more than once to `COUNT` or `ALL`.

## A useful diagnostic sequence

1. Confirm canonical metadata names and supported selectors.
2. Run `validate_anchor_inputs()` without concepts.
3. Inspect invalid and unexpectedly expanded rows from `define_window()`.
4. Test one variable against a small in-memory concept subset.
5. Run the production parquet or DuckDB source.
6. Read long output first; reshape to wide only after confirming result cardinality.
